---
title: "WES Report"
author: "Stephen Kelly"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    css: styles.css
    keep_md: yes
    number_sections: true
    toc: true
    toc_float: true
params:
  run_analysis_output_dir: run_analysis_output
  sns_wes_coverage_analysis_dir: sns-wes-coverage-analysis
  project_ID_file: project_ID.txt
  results_ID_file: results_ID.txt
---
<!-- Setup the R code to be used in R Markdown generation throughout the report -->
```{r setup, include=FALSE} 
# {.tabset} # .tabset-fade .tabset-pills

# ~~~~~ SETTINGS ~~~~~ #
knitr::opts_chunk$set(echo = FALSE)

# ~~~~~ LOAD PACKAGES ~~~~~ #
library("knitr")
library("ggplot2")
library("reshape2")
library("data.table")
library("DT")
# ~~~~~ CUSTOM FUNCTIONS ~~~~~ #
mycat <- function(text){
    # function for formatting text in the report
    cat(gsub(pattern = "\n", replacement = "  \n", x = text))
}

file_scan <- function(my_file){
    # function for reading info from a text file
    file_contents <- scan(my_file, what="", sep="\n")
    return(file_contents)
}

get_file_content <- function(my_file){
    # function to return file contents or NA if file doesn't exit
    file_contents <- "NA"
    if (file.exists(my_file)) {
        if (! identical(file_scan(my_file), character(0))) {
            file_contents <- file_scan(my_file)
        }
    }
    return(file_contents)
}

reindex_rownames <- function(df){
    # reindex the row names after subsetting
    if(nrow(df) > 0) rownames(df) <- seq(nrow(df))
    return(df)
}

chrom_regions2df <- function(regions){
    # split the regions into chrom coordinates for BED files
    # regions <- c("chr1:236998847-236998987", "chr1:237001714-237001899")
    regions_df <- as.data.frame(do.call(rbind, strsplit(regions, ':')))
    regions_df <- cbind(regions_df[1],
                        as.data.frame(do.call(rbind, strsplit(as.character(regions_df$V2), '-'))))
    colnames(regions_df) <- c("chrom", "start", "stop")
    return(regions_df)
}

# ~~~~~ LOAD REPORT DATA ~~~~~ #
report_dir <- getwd() # setwd(report_dir)
sns_wes_coverage_analysis_dir <- params$sns_wes_coverage_analysis_dir
run_analysis_output_dir <- params$run_analysis_output_dir
project_ID_file <- params$project_ID_file
results_ID_file <- params$results_ID_file

project_ID <- get_file_content(project_ID_file)
results_ID <- get_file_content(results_ID_file)

summary_combined_table_file <- file.path(run_analysis_output_dir, "summary-combined.wes.csv")

summary_combined_table <- read.delim(file = summary_combined_table_file, sep = ',')
setnames(x = summary_combined_table, old = c("X.SAMPLE"), new = c("sample"))

save.image(file="load_report_data.Rdata",compress = TRUE)
```

```{r, samples_table}
samples_df <- summary_combined_table[, "sample", drop = FALSE]
```

```{r, mapping_table}
mapping_table <- summary_combined_table[, "sample", drop = FALSE]
# mapping_table["avg_raw_reads"] <- as.numeric(summary_combined_table[["R1.RAW.READS"]]) + as.numeric(summary_combined_table[["R2.RAW.READS"]]) / 2
mapping_table["mapped"] <- summary_combined_table[["MAPPED.READS"]]
mapping_table["deduplicated"] <- summary_combined_table[["DEDUPLICATED.READS"]]

mapping_table <- reshape2::melt(mapping_table,id.vars="sample",variable.name="type",value.name="reads")

mapping_table[["reads"]] <- mapping_table[["reads"]] / 1e6

mapping_table[["type"]] <- factor(x = mapping_table[["type"]], levels = sort(unique(as.character(mapping_table[["type"]]), decreasing = TRUE)))

mapping_plot <- ggplot(data = mapping_table, aes(x = sample, y = reads, fill = type)) + geom_bar(stat="identity", position = "dodge") + ggtitle("Sample Read Mapping") + coord_flip() + ylab("reads (millions)")  

```

```{r, mean_coverages}
mean_coverages_table <- summary_combined_table[c("sample", "MEAN.COVERAGE", "MEDIAN.COVERAGE")]
setnames(mean_coverages_table, old = c("MEAN.COVERAGE", "MEDIAN.COVERAGE"), new = c("mean", "median"))
mean_coverages_table <- reshape2::melt(mean_coverages_table, id.vars="sample", variable.name="type",value.name="coverage")

mean_coverages_plot <- ggplot(data = mean_coverages_table, aes(x = sample, y = coverage, fill = type)) + geom_bar(stat="identity", position = "dodge") + ggtitle("Coverage Per Sample") + coord_flip()
```

```{r, coverage_cutoffs}
# colnames of the coverage cutoff columns
coverage_above_cols <- c("X._bases_above_10", "X._bases_above_50", "X._bases_above_100", "X._bases_above_500")

# make a table for just the coverage cutoff data
coverage_cutoff_table <- reshape2::melt(summary_combined_table[c("sample", coverage_above_cols)],id.vars="sample",variable.name="coverage_cutoff",value.name="percent")

# fix the values in the cutoff column
coverage_cutoff_table[["coverage_cutoff"]] <- gsub(pattern = 'X._bases_above_', replacement = '', x = as.character(coverage_cutoff_table[["coverage_cutoff"]]))
# reorder the factor levels for the plot
coverage_cutoff_table[["coverage_cutoff"]] <- factor(coverage_cutoff_table[["coverage_cutoff"]], levels = sort(unique(as.numeric(coverage_cutoff_table[["coverage_cutoff"]]))))

# make the plot
coverage_cutoff_plot <- ggplot(data = coverage_cutoff_table, aes(x = sample, y = percent, fill = coverage_cutoff)) + geom_bar(stat="identity", position = "dodge") + ggtitle("Percent of Bases Above Coverage Cutoff") + coord_flip()

```

```{r, low_coverage_regions}
low_cov_cols <- c("gene", "chrom", "start", "end", "width", "distancetoFeature", "biotype")
## # # # # 
average_coverage_per_sample_table <- read.delim(file.path(sns_wes_coverage_analysis_dir, "average_coverage_per_sample.tsv"))
average_coverage_per_region_table <- read.delim(file.path(sns_wes_coverage_analysis_dir, "average_coverage_per_region.tsv"), header = FALSE)
# average_coverage_per_region_regions <- chrom_regions2df(as.character(average_coverage_per_region_table[[1]]))

average_coverage_per_region_plot <- ggplot(data=average_coverage_per_region_table, aes(average_coverage_per_region_table$V2)) + geom_histogram(binwidth = 50) + ggtitle("Average Coverage per Region Histogram") + ylab("number of regions") + xlab("average coverage value")

## # # # # 

regions_coverage_below_50_annotation_table <- read.delim(file.path(sns_wes_coverage_analysis_dir, "regions_coverage_below_50_annotation.tsv"))


setnames(x = regions_coverage_below_50_annotation_table, old = c("external_gene_name", "seqnames", "gene_biotype"), new = c("gene", "chrom", "biotype"))

regions_coverage_below_50_annotation_table <- regions_coverage_below_50_annotation_table[order(regions_coverage_below_50_annotation_table[["gene"]]),]
regions_coverage_below_50_annotation_table <- reindex_rownames(regions_coverage_below_50_annotation_table)




## # # # # 
regions_0_coverage_annotation_table <- read.delim(file.path(sns_wes_coverage_analysis_dir, "regions_with_coverage_0_annotation.tsv"))
setnames(x = regions_0_coverage_annotation_table, old = c("external_gene_name", "seqnames", "gene_biotype"), new = c("gene", "chrom", "biotype"))
regions_0_coverage_annotation_table <- regions_0_coverage_annotation_table[order(regions_0_coverage_annotation_table[["gene"]]),]
regions_0_coverage_annotation_table <- reindex_rownames(regions_0_coverage_annotation_table)


```

```{r, variant_calling}
haplotype_caller_table_og <- read.delim(file.path(run_analysis_output_dir,"summary.VCF-GATK-HC-annot.csv"), sep = ',')
setnames(x = haplotype_caller_table_og, old = c("X.SAMPLE"), new = c("sample"))
haplotype_caller_table <- haplotype_caller_table_og
haplotype_caller_table <- reshape2::melt(haplotype_caller_table, id.vars="sample",variable.name="type",value.name="mutations")
haplotype_caller_table[["type"]] <- gsub(pattern = '.muts', replacement = '', x = as.character(haplotype_caller_table[["type"]]))
haplotype_caller_plot <- ggplot(data = haplotype_caller_table, aes(x = sample, y = mutations, fill = type)) + geom_bar(stat="identity", position = "dodge") + ggtitle("Number of Mutations per Sample") + coord_flip()
# # # # # 
lo_freq_table_og <- read.delim(file.path(run_analysis_output_dir,"summary.VCF-LoFreq-annot.csv"), sep = ',')
setnames(x = lo_freq_table_og, old = c("X.SAMPLE"), new = c("sample"))
lo_freq_table <- lo_freq_table_og
lo_freq_table <- reshape2::melt(lo_freq_table, id.vars="sample",variable.name="type",value.name="mutations")
lo_freq_table[["type"]] <- gsub(pattern = '.muts', replacement = '', x = as.character(lo_freq_table[["type"]]))
lo_freq_plot <- ggplot(data = lo_freq_table, aes(x = sample, y = mutations, fill = type)) + geom_bar(stat="identity", position = "dodge") + ggtitle("Number of Mutations per Sample") + coord_flip()

```

# Overview

Project ID:

`r project_ID`

Results ID:

`r results_ID`

```{r, results='asis'}
kable(samples_df, row.names = TRUE, caption = "Samples included in the analysis") # , align = c("c")
# datatable(samples_df, caption = "Samples included in the analysis")

```

## Mapping

```{r, fig.height=12, fig.width=8}
print(mapping_plot)
```

## Coverage Per Sample {.tabset .tabset-pills}

### Mean

```{r}
print(mean_coverages_plot)
```

### Per Cutoff

```{r, fig.height=12, fig.width=8}
print(coverage_cutoff_plot)
```

# Target Regions

## Coverage Per Region

```{r}
print(average_coverage_per_region_plot)
```

## Regions With Low Coverage (<50)

```{r}
datatable(regions_coverage_below_50_annotation_table[low_cov_cols])
```

## Regions with 0 Coverage

```{r}
datatable(regions_0_coverage_annotation_table[low_cov_cols])
```

# Variant Calling

## Haplotype Calling {.tabset .tabset-pills}

### Plot

```{r}
print(haplotype_caller_plot)
```

### Table

```{r}
datatable(haplotype_caller_table_og)
```

## Lo Freq {.tabset .tabset-pills}

### Plot 

```{r}
print(lo_freq_plot)
```

### Table

```{r}
datatable(lo_freq_table_og)
```


# System Information {.tabset .tabset-pills}

## Hide

## Show

```{r}
# system info
system("uname -srv", intern = TRUE)

# dir
system('pwd',intern=T)

# repo info
system('git remote -v',intern=T)
system('printf "%s\t%s" "$(git rev-parse --abbrev-ref HEAD)" "$(git rev-parse HEAD)"',intern=T)

# date time
system("date", intern = TRUE)

# R system info, packages, etc
sessionInfo()

# save current session
save.image(file="final_report_data.Rdata",compress = TRUE)
```
